---
title: "Homework 01"
---


## TidyTuesday Section

Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

### Research Question: 

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/movies.csv')
shows <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/shows.csv')
```
```{r}
movies
shows
```
```{r}
movies_month <- movies |>
  mutate(month = month(release_date, label = TRUE))

shows_month <- shows |>
  mutate(month = month(release_date, label = TRUE))

ggplot(movies_month, aes(x = month, y = views)) +
  geom_boxplot(fill = "blue", outlier.alpha = 0.3) +
  scale_y_log10(labels = scales::comma) +
  labs(title = "Netflix Movies: Views by Release Month",
       x = "Release Month",
       y = "Views (log scale)") +
  theme_minimal()

ggplot(shows_month, aes(x = month, y = views)) +
  geom_boxplot(fill = "red", outlier.alpha = 0.3) +
  scale_y_log10(labels = scales::comma) +
  labs(title = "Netflix Shows: Views by Release Month",
       x = "Release Month",
       y = "Views (log scale)") +
  theme_minimal()
```
```{r}
movies$type <- "Movies"
shows$type <- "Shows"

combined <- dplyr::bind_rows(movies, shows) |>
  mutate(month = factor(format(release_date, "%b"), levels = month.abb))

ggplot(combined, aes(x = month, y = views, fill = type)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_y_log10(labels = scales::comma) +
  facet_wrap(~ type) +
  labs(title = "Netflix Movies vs Shows: Views by Release Month",
       x = "Release Month", y = "Views (log scale)") +
  theme_minimal()
```


Based on the two boxplot graph above, we can see that overall, movies tend to have slightly higher median views than shows, which means that films gnerally attract larger audiences than shows. 
For movies, there is some seasonal variations that movie released in July and November show relaitively higher views, which means likely relate to sumer and holiday breaks when people have more free time to watch films. 
In contract, shows display a more stable pattern throughout the year. 
Both movies and shows have notable outliers, which means that there are some films that hit exceptional viewership in regard of the release time. 

In conclusion: Movies are more sensitivezen to seasonal effects, while shows have more steady audence engagement all ear round. 


